---
title: "Network of Taylor Swift's Album Credits"
author: "Eduard De Guzman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

For this project, we have created a mini-version of the networks shown in [this](https://sonic-edgehog.github.io/general/2021/04/04/ts-blog-001/). Since `plotly` in `Python` does not have the ready to made interactive version of the `networkx`counterpart, we have used the platform for that. But before graphing, we have saved the necessary data frames of nodes and links for production and lyric credit.

For details regarding the library, please see this [documentation](https://www.rdocumentation.org/packages/networkD3/versions/0.4/topics/forceNetwork)

## Exploration

We used `magrittr` to save the file into a stand-alone html and `networkD3`, a library with integration of styles from `JS` and `R`.

```{r introduction}
library(networkD3)
library(magrittr)


ts_prod_all_link = read.csv('ts_network_prod_all.csv')
ts_prod_all_node = read.csv('ts_nodes_prod_all.csv')
```

Here are the contents of the following tables:

```{r}
head(ts_prod_all_link)
```

The first two columns are the pairs that worked together in producing the song and the last column is the number of occurrences.

```{r}
head(ts_prod_all_node)
```
For the dataset, `size` is the number of works a producer has solely done plus one while `album_work` is the eras in which a producer has mainly worked only.

## Data Manipulation

Since the package `networkD3` does not work on pairs which are non-numeric, we have used the left join for both datasets. To highlight the graphs which **Jack Antonoff** has worked on, we modified the links.

```{r}
ts_prod_all_link = merge(x=ts_prod_all_link,y=ts_prod_all_node[c("name_a","source")],by="name_a",all.x=TRUE)

ts_prod_all_link = merge(x=ts_prod_all_link,y=ts_prod_all_node[c("name_b","sink")],by="name_b",all.x=TRUE)

jack_number = which(ts_prod_all_node == "Jack Antonoff", arr = TRUE)[1] - 1
jacklink = which(ts_prod_all_link == jack_number, arr = TRUE)[, 1]

jack_color = ifelse(1:nrow(ts_prod_all_link) %in% jacklink, "#bf3eff", "#666")
```

## Plot of Producers Network.

Then, we plot the graphs themselves. We have specified the colors for each category in `album_work` column and modified the sizes according to a function.

```{r}
graph_prod = forceNetwork(Links = ts_prod_all_link, Nodes = ts_prod_all_node,Nodesize="size",
             Source = "source", Target = "sink",colourScale=JS("d3.scaleOrdinal(['#383867', '#b63e36', '#33431e', '#b74a70','#a36629']);"),
             Value = "value", NodeID = "name_a",radiusCalculation = "Math.sqrt(d.nodesize)+7",
             Group = "album_work", opacity = 0.8,
             fontFamily = "sans-serif",
             fontSize = 20,legend=TRUE,zoom = TRUE,bounded = TRUE,
             opacityNoHover = 0.1,linkColour = jack_color)
graph_prod
```

It seems that one, everyone except the pair who worked on `A Perfectly Good Heart` in the first album are connected. Two, **Jack Antonoff** has connected the pop and alternative musicians with his participation both in `1989` and `folklore`.

```{r}
graph_prod %>% saveNetwork(file = 'ts_graph_prod_all.html',selfcontained = TRUE)
```

To save the network as a stand-alone, I used the function for the package.

## Replication for Co-Writers

First, we read the tables.

```{r}
ts_write_all_link = read.csv('ts_network_write_all.csv')
ts_write_all_node = read.csv('ts_nodes_write_all.csv')
```

Then, we joined them together.

```{r}
ts_write_all_link = merge(x=ts_write_all_link,y=ts_write_all_node[c("name_a","source")],by="name_a",all.x=TRUE)

ts_write_all_link = merge(x=ts_write_all_link,y=ts_write_all_node[c("name_b","sink")],by="name_b",all.x=TRUE)

jack_number = which(ts_write_all_node == "Jack Antonoff", arr = TRUE)[1] - 1
jacklink = which(ts_write_all_link == jack_number, arr = TRUE)[, 1]

jack_color = ifelse(1:nrow(ts_write_all_link) %in% jacklink, "#bf3eff", "#666")
```

After that, we graphed them.

```{r}
graph_write = forceNetwork(Links = ts_write_all_link, Nodes = ts_write_all_node,Nodesize="size",
             Source = "source", Target = "sink",colourScale=JS("d3.scaleOrdinal(['#b63e36','#33431e','#b74a70','#a36629']);"),
             Value = "value", NodeID = "name_a",radiusCalculation = "Math.sqrt(d.nodesize)+7",
             Group = "album_work", opacity = 0.8,fontFamily = "sans-serif",
             fontSize = 20, legend=TRUE,zoom = TRUE,bounded = TRUE,
             opacityNoHover = 0.1,linkColour = jack_color)
graph_write
```

Among Taylor's co-writers, there are two main sub-graphs. Namely, the one who worked for `1989` and `reputation` while the another one worked for albums `post-Red`. In this case, **Jack Antonoff** is Taylor's main co-writer from albums `post-Red`. Furthemore, **Max Martin** and **Shellback** have written together the most yet the former and Taylor Swift only worked together once, which is `All You Had To Do Was Stay` from `1989`.

```{r}
graph_write %>% saveNetwork(file = 'ts_graph_write_all.html',selfcontained = TRUE)
```

Finally, we saved them.

## Conclusion

We concluded that **Nathan Chapman**, **Aaron Dessner**, and **Jack Antonoff** are Taylor's main co-workers. Also, most writers and producers (excluding Jack Antonoff and the duo Max Martin-Shellback) in `1989` and `reputation` worked only during those album. Lastly, **Aaron Dessner** likes to work alone or solely with Taylor.